---
title: "Guide to aplicitation of LBPA in data-limited fisheries situations"
subtitle: "Length Based Pseudo-Cohort Analysis"
author: "Canales, C., Punt, AE., Mardones, M."
date: "`r Sys.Date()`"
output: pdf_document
toc: TRUE
---

```{r setup, include=FALSE, warning=FALSE, error=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(patchwork)
library(knitr)
```

\pagebreak

# Introduction

This guide contains functions to compile and run the Length-based PseudoCohort Analysis (LBPA) fisheries stock assessment method in data limited conditions

This model is an estimation model by fitting to length composition data to estimate annual fishing mortality, annual recruitment and spawning potential ratio (SPR)

LBPA was developed for data-limited fisheries, where few data are available other than a representative sample of the size structure of the vulnerable portion of the population (i.e., length composition data from the catch) and an understanding of the life history of the species.

# Basic functionality

The LBPA model is built based on a C ++ language programmed in ADMB, and below we provide the steps to compile and build from a simple code implemented in R that had next step.

To do this, you must have installed the ADMB version (12 or newer). Can download in this site: <http://www.admb-project.org/>

The LBPA model and data example can be obtained from <https://github.com/CristianCanales/LBPA>


### Bug Reports
Alert to any bugs or issues by using GitHub. Suggestions and comments for additional features are welcome and we can discuss via email at

# Step 1

## Set working directory

In this folder you need put two files, .tpl and .dat to run LBPA estimation model 

```{r echo=TRUE, eval=FALSE}
rm(list=ls(all=TRUE))  # erasure all objects
setwd('~/')
```

Once you have defined your working directory, you can take a look and check the contains files.

```{r eval=TRUE, echo=FALSE}
dir()
```

Is necessary to get a function to can read report. This function is hold in the same folder than .tpl and .dat.


```{r echo=TRUE, eval=FALSE}
source('~/read.admb.R')
```

# Step 2

## Run model

Now we compile it in ADMB set. If you have read it, dont need run again:

```{r, warning=FALSE, eval=FALSE, echo=TRUE}
system('~/admb-12.2/admb LBPA')
```

Now you need run the model with console. The *system* function could run *.tpl* code from R.

```{r, warning=FALSE, eval=FALSE, echo=TRUE}
system('./LBPA')
```

\pagebreak

# Step 3

## Read Report

Once run our LBPA model, we can read report with *read.rep()* function.
```{r}
data <-read.rep('LBPA.rep')
```

In data example we have a set data from a clam  population from Bahía de Ancud, in southern Chile. You can change the data relative to your stock in the *.dat* hosted in the github repository

Now, we can read data in our .rep file.

```{r echo=FALSE}
names(data)
```

We need assign variables to get different plots;

```{r}
BinLen <- data$Length
ObsFre <- data$Observed_frequencies  
PredFre <- data$Predicted_frequency 
CatchLFre <- data$Catch_length_frequency 
ProbLen <- data$Probability_of_length
```

Lets plot size structure from data

```{r, echo=TRUE, fig.align="center", fig.cap= "Set of Length data with model adjust"}
par(mfrow=c(3,3),mar=c(3,4,1,1)+1)
for(i in 1:9){
  plot(BinLen, ObsFre[i,],type="h", xlab="mm.", ylab="Lenght Density") 
  lines(BinLen, PredFre, type="l", col=2)
 }

```

\pagebreak

we can do a plot with the sum of all length set data available;

```{r  fig.cap= "Sum of Length data with model adjust"}
plot(BinLen, ObsFre[7,],type="s", xlab="mm.", ylab="Lenght Density")
lines(BinLen, PredFre, type="l", col=2, lwd=4)
legend("topleft", c("Length Predicted", "Length Observed"), col=c(2, 1), lwd =c(1.5, 1.5), lty=c(1,1), bty="n", cex=0.8)
```

\pagebreak
# Step 4

## Ploting outputs LBPA

Now we plot diferent outputs, as lengths, adjust models, fishing mortality, selectivity, maturity ogive and SPR.


### Plot of estimates parameters

```{r eval=TRUE, echo=FALSE}
age    <- seq(1,10)
Length <- (data$Age_Length_SE_N_Catch_Select_F[,2])
SE     <- (data$Age_Length_SE_N_Catch_Select_F[,3])
Abun   <- (data$Age_Length_SE_N_Catch_Select_F[,4])
Catch  <- (data$Age_Length_SE_N_Catch_Select_F[,5])
Selec  <- (data$Age_Length_SE_N_Catch_Select_F[,6])
Fm     <- (data$Age_Length_SE_N_Catch_Select_F[,7])
```


```{r fig.cap="Parameters used in LBPA estimation model"}
p <- ggplot(data=NULL, aes(x=age))+
  geom_line(aes(y = Length,  colour=2))+
  theme_bw()
q <- ggplot(data=NULL, aes(x=age))+
  geom_line(aes(y = SE,  colour=3))+
  theme_bw()
s <- ggplot(data=NULL, aes(x=age))+
  geom_line(aes(y = Fm,  colour = 4))+
  theme_bw()
t <- ggplot(data=NULL, aes(x=age))+
  geom_line(aes(y = Abun,  colour = 5))+
  theme_bw()
o <- ggplot(data=NULL, aes(x=age))+
  geom_line(aes(y = Catch,  colour = 6))+
  theme_bw()
v <- ggplot(data=NULL, aes(x=age))+
  geom_line(aes(y = Selec,  colour = 7))+
  theme_bw()
p/q|s/t|o/v
```

\pagebreak

###  Length current, objetctive and expected

```{r}
L1 <- (data$Length_frequency_of_exploitable_population[1,])
L2 <- (data$Length_frequency_of_exploitable_population[2,])
L3 <- (data$Length_frequency_of_exploitable_population[3,])
```


```{r echo=FALSE, fig.cap="Set of Length estimated in LBPA" }
L123 <- ggplot(data=NULL, aes(x=BinLen))+
  geom_line(aes(y = L1, colour = 'Unfished'))+
  geom_line(aes(y = L2, colour = 'Target'))+
  geom_line(aes(y = L3, colour = 'Current'))+
  scale_color_manual(name = '',
                     values = c('red1', 'gold1', 'green'),
                     limits = c('Unfished', 'Target', 'Current'),
                     breaks = c('Unfished', 'Target', 'Current')) +
  theme(legend.position = 'bottom') + ylab('Frecuency') + xlab('Length (mm.)') +
  theme_bw()
L123
```

### Table Parameters Models used in LBPA estimation model;

```{r error=FALSE, warning=FALSE}
tabla <- matrix(ncol=1, round(data$F_L50_slope_a0_cv_Lr_Ftar, 3))
rownames(tabla) <- c("F", "L50", "Slope", "a0", "cv", "Lr", "F_Target")
kable(tabla)
```


